---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #1
## Blake, Amaya, Brietta, Simeon

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

```{r}
library(classdata) # Simeon
library(ggplot2) # Blake
```

1. inspect the first few lines of the data set:

what variables are there? of what type are the variables? what does each variable mean? what do we expect their data range to be?


The 16 variables are: 
```{r}
str(ames) #blake
colnames(ames)#blake
head(ames)#brietta
```

#brietta 

Numerical - 7
Strings/Character - 5
Factor - 3
Date - 1

  Parcel ID - Numerical
    general number given for the listing - range could be any number following numerically after the previous one - incremented
  Address - String   
    could be any combination of numbers and letters based on the address
  Style - Factor  
    descriptor for the house - range could be anything descriptive for the house or apartment
  Occupancy - Factor 
    same as above but describes how many people and what type of building it is - ex condo
  Sale Date - Date  
    range is any date that the thing was purchased - range is any logical date
  Sale Price - Numerical
    no range for sale proce
  Multi Sale -  String
    if there was more than one building purchaed - NA or other input
  YearBuilt - Numerical   
    has to be wihtin a range of years that is logical
  Acres - Numerical
    any number - has to make sense 
  TotalLivingArea (sf) - Numerical
    same as acres
  Bedrooms - Numerical
    whole number - also has to be a logical number
  FinishedBsmtArea (sf) - Numerical
    area of the basement - should be around the same as living area
  LotArea(sf) - Numerical
    same as basement - area of loft - should be around the same area as the living area
  AC - String
    yes or no for AC
  FirePlace - String 
    yes or no for fireplace
  Neighborhood - Factor        
    descriptor for describing the area - no range
    
2. is there a variable of special interest or focus?
# Simeon
  The variable of interest is Sale Price
  
3. start the exploration with the main variable:

  - what is the range of the variable? draw a histogram for a numeric variable or a bar chart, if the variable is categorical. what is the general pattern? is there anything odd?


  - follow-up on oddities: see 4


range(ames$`Sale Price`, na.rm = TRUE) #Amaya
hist(ames$`Sale Price`, main = "Histogram of Sale Prices", xlab = "Sale Price", breaks = 30) #Amaya



  - follow-up on oddities: see 4
 
The range of the Sale Price is:
0 - 20500000

(Range code shown twice) 
```{r}

  range(ames$`Sale Price`, na.rm = TRUE) #Amaya
  range(ames$`Sale Price`) #blake
```

Histogram


```{r}
  hist(ames$`Sale Price`, main = "Histogram of Sale Prices", xlab = "Sale Price", breaks = 30)
#Amaya
```


  

  - follow-up on oddities: see 4\

    The range of the Sale Price is:

```{r}
    range(ames$`Sale Price`)
    #blake
```


4. `pick a variable that might be related to the main variable.


4. pick a variable that might be related to the main variable.
#Simeon

```{r}
range(ames$`Sale Date`)
  
```

  - what is the relationship to the main variable? plot a scatterplot, boxplot or facetted barcharts (dependening on the types of variables involved). Describe overall pattern, does this variable describe any oddities discovered in 3? Identify/follow-up on any oddities.

  
4. `pick a variable that might be related to the main variable.
```{r} 
#Amaya
# Range of Bedrooms
range(ames$Bedrooms, na.rm = TRUE)

```{r}
ames$`Sale Date` <- as.Date(ames$`Sale Date`)
plot(ames$`Sale Date`, ames$`Sale Price`, 
     main = "Scatter Plot of Sale Date vs. Sale Price", 
     xlab = "Sale Date", 
     ylab = "Sale Price", 
     col = "blue", 
     pch = 19)
```


4. pick a variable that might be related to the main variable. 
#Brietta

  - what is the range of that variable? plot. describe the pattern.
  
```{r}
  mean(ames$"TotalLivingArea",na.rm = TRUE)
  range(ames$"TotalLivingArea",na.rm = TRUE)
```

the range is 0 - 6007


  - what is the relationship to the main variable? plot a scatterplot, boxplot or facetted barcharts (dependening on the types of variables involved). Describe overall pattern, does this variable describe any oddities discovered in 3? Identify/follow-up on any oddities.
```{r}

#Bar chart for number of Bedrooms
barplot(table(ames$Bedrooms), main = "Distribution of Bedrooms", xlab = "Number of Bedrooms", ylab = "Frequency")

#Boxplot of Sale Price vs. Bedrooms
boxplot(ames$`Sale Price` ~ ames$Bedrooms, main = "Sale Price by Number of Bedrooms", xlab = "Number of Bedrooms", ylab = "Sale Price")

#Amaya 
```

Bedroom Range: The number of bedrooms in the dataset ranges from X to Y.
Bar Chart Pattern: The majority of homes have X bedrooms, with fewer homes having more than Y bedrooms.
Boxplot Observation: The number of bedrooms generally correlates with higher sale prices, but there is some variability, with homes having the same number of bedrooms displaying a wide range of sale prices.

4. pick a variable that might be related to the main variable.
#Blake
Acres

```{r}
#The range of the acreage is shown in the boxplot that follows:
ggplot(ames, aes(x=Acres))+
  geom_boxplot()
```

